---
title: "Using *clusso* with Case-Control Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to clusso II}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 5, fig.width = 8)
```

#Introduction

This package is still under development. Please [report any bugs](https://github.com/mkamenet3/clusso/issues).

Messages from `clusso` have been suppressed for this tutorial, but there are many messages generated that we hope the user will find helpful. Feedback is welcome [here](https://github.com/mkamenet3/clusso/issues).

Load `clusso` package:

```{r, message=FALSE, warning=FALSE}
library("clusso")

```

If you haven't already, you can install `clusso` directly from GitHub:

```{r, eval=FALSE}
library("devtools")
devtools::install_github("mkamenet3/clusso")
```



## Using `clusso` with case-control data 

In this example, we will use simulated case-control data of across 208 prefects (municipalities) in Japan across 5 time periods.

### Prepare Data

We load four data sets that come with the `clusso` package:

```{r}
#load data
data("ccsjbc")
data("utmJapan")

```


These data sets contain:

1. `ccsjbc`: data set containing a unique identifier for each prefect centroid (*id*), period of observation (categorical variable with 5 levels, *period*), number of cases per each geographic unit (*numcases*), the total number of cases and controls from each prefect (*n*). These data have been simulated.
1. `utmJapan`: data set containing a unique identifier for each prefect centroid (*id*), x-coordinate in UTM (*utmx*), and y-coordinate in UTM (*utmy*).


To explore the 4 data sets, we apply the `head()` function to each of the data sets.



```{r}
#inspect
head(utmJapan)
head(ccsjbc)
```


As a data-cleaning step, we first convert `period` to a factor in the data frame `jbc`:

```{r}
ccsjbc$period <- as.factor(ccsjbc$period)
ccsjbc$id <- as.factor(ccsjbc$id)

```



### Set global parameters

When using `clusso`, there are certain global parameters that need to be set.

We take the easting and northing coordinates from the `utmJapan` data set, set the easting coordinate, `utmx`, to `xx` and the northing coordinate, `utmy`, to `yy`. Each is divided by 1000 to change the scale from meters to kilometers.


```{r}
xx <- utmJapan$utmx/1000
yy <- utmJapan$utmy/1000
```

Below, we set the maximum radius for a cluster to 20 km (`rMax=20`) based on scientific knowledge of the area and our definition of a meaningful cluster, set the number of unique time periods to 5 (`Time=5`). 


```{r}
rMax <- 20 
Time <- 5
```


### Fitting using `clusso`

The argument `model` is the type of model to be used. In this case-control study example, we specify the binomial model.

The output of `clusso()` is assigned to the object `resreal`. `resreal` is a large list of lists. The recommended way to explore the results is to explore the `names(resreal)` and select each sub-list of interest.

```{r, warning=FALSE}
head(ccsjbc)

system.time(resrealccs <- clusso(df=ccsjbc, expected = n, observed = numcases,
                                 timeperiod = period,id=id,
                                 covars=FALSE, x= xx,y = yy, 
                                 rMax =  rMax, utm=TRUE, 
                                 analysis="both", model="binomial",
                                 maxclust=15, collapsetime = FALSE))

```


```{r, warning=FALSE}
clussopretty(resrealccs, analysis="both", model="binomial",clusteridentify=FALSE)
```


## Case-Control Data with Covariates

We continue with the case-control data from Japan. We will again checkout the `head()` of the dataframe, `ccsjbc`. 


```{r}
head(ccsjbc)
```


There are 5 additional covariates, `covar1` - `covar5` that we will want to adjust for as unpenalized terms.


We now set the argument `covars=TRUE` to tell `clusso()` to identify the covariates and to include them in as unpenalized terms.



```{r, warning=FALSE}
system.time(resccscovars<- clusso(ccsjbc, expected = n, observed=numcases,timeperiod = period, id=id,
                                  covars=TRUE, x= xx,y = yy, rMax =  rMax, utm=TRUE,
                                  analysis="both", model="binomial",maxclust=11, collapsetime=FALSE))

```



```{r, warning=FALSE}
clussopretty(resccscovars, analysis="both", model="binomial",clusteridentify = FALSE)
```


We can also extract the coefficients for the unpenalized terms and undo the log-link by exponentiating: 

```{r}
out <- clussopretty(resccscovars, analysis="both", model="poisson",covars=TRUE)
out$table.coefs
out$table.expcoefs
out$table.clusters
```




































