---
title: "Introduction to *clusso*"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to clusso I}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 5, fig.width = 8, message = TRUE)
```

#Introduction

This package is still under development. Please [report any bugs](https://github.com/mkamenet3/clusso/issues).

Messages from `clusso` have been suppressed for this tutorial, but there are many messages generated that we hope the user will find helpful. Feedback is welcome [here](https://github.com/mkamenet3/clusso/issues).

Load the `clusso` library:

```{r, message=FALSE, warning=FALSE}
library("clusso")

```

If you haven't already, you can install `clusso` directly from GitHub:

```{r, eval=FALSE}
library("devtools")
devtools::install_github("mkamenet3/clusso")
```


## Using `clusso` (Breast Cancer in Japan Example)

In this example, we will use simulated data of breast cancer incidence counts across 208 prefects (municipalities) in Japan across 5 time periods.

### Prepare Data

We load four data sets that come with the `clusso` package:


```{r}
#load data
data("jbc")
data("utmJapan")

```


To explore the 4 data sets, we apply the `head()` function to each of the data sets.

1. `utmJapan`: data set containing a unique identifier for each prefect centroid (*id*), x-coordinate in UTM (*utmx*), and y-coordinate in UTM (*utmy*).
1. `jbc`: data set containing a unique identifier for each centroid (*id*), period of observation (categorical variable with 5 levels, *period*), death count (*death*), expected death count (*expdeath*). These data are simulated (with some noise) based on observed data.



```{r}
#inspect
head(utmJapan)
head(jbc)

```

We first convert `period` to a factor in the data frame `jbc`:

```{r}
jbc$period <- as.factor(jbc$period)
```



### Set global parameters

When using `clusso`, there are global parameters that need to be set first.


We take the easting and northing coordinates from the `utmJapan` data set, set the easting coordinate, `utmx`, to `x` and the northing coordinate, `utmy`, to `y`. Each is divided by 1000 to change the scale from meters to kilometers.


```{r}
x <- utmJapan$utmx/1000
y <- utmJapan$utmy/1000
```

Below, we set the maximum radius for a cluster to 20 km (`rMax=20`) based on scientific knowledge of the area and our definition of a meaningful cluster. We set the number of unique time periods to 5 (`Time=5`), and set a floor for overdispersion to avoid estimating underdispersion (`overdispfloor=TRUE`). 


```{r}
rMax <- 20 
Time <- 5
overdispfloor <- TRUE

```

If you do not have expected counts readily available, they can be calculated using the auxiliary function `calcexpected()`, from the `clusso` package.

As an example, we generate some data and apply `calcexpected()`. The output is a dataframe with an additional variable, `expected`, which contains the expected counts:

```{r}
set.seed(2)
period <- rep(c("1","2"), times=5)
observed <- MASS::rnegbin(n = length(period), mu=20, theta=2)
pop <-  MASS::rnegbin(n = length(period), mu=200, theta=2)
ids <- rep(1:5, each=2)

calcexpected(observed, pop, period, ids)
```

In this data example, we already have expected counts and will use the dataframe `jbc`.

### Fitting using `clusso`


To perform the regularized spatial and spatio-temporal cluster detection using the Lasso penalty, we use the function `clusso()`. In the first argument, we set the name of the dataframe, `df=jbc`. In the second argument we set the expected counts equal to the name of expected counts in the `jbc` dataframe (`expected = expdeath`). In the third argument, we set the observed counts equal to the name of the observed counts in the `jbc` dataframe (`observed = death`). In the fourth argument, we set the time period equal to the name of the time periods in the `jbc` dataframe (`timeperiod = period`). This variable has already been converted to a factor in the data-cleaning step above (`jbc$period <- as.factor(jbc$period)`). In this example, we do not want to include other covariates as unpenalized terms in the model so we set `covars = FALSE`. 

The next two arguments are `x` and `y`, which are the UTM easting and northing coordinates, scaled to be in kilometers. The next argument is the maximum radius (in km) for the potential clusters. We set `rMax = rMax`, which we defined above as 20km. 
The argument `utm` specifies that our coordinates are indeed in UTM coordinates (`utm = TRUE`). For latitude-longitude coordinates, we would set `utm = FALSE`. The is `analysis` argument defines the type of analysis to be performed. Options include `space`, `spacetime`, or `both`. Here, we set  `analysis` equal to "both", which will perform both the spatial and spatio-temporal analyses. The argument `model` is the type of model to be used. Currently (2020-09-14), either the Poisson or binomial models can be specified. In this example, we specify the Poisson model. The final required argument is `maxclust`. This is the maximum number of actual clusters expected in the given area. This is based on scientific knowledge of the area. For the Japanese breast cancer data, we set `maxclust=11`. If the number of identified clusters exceeds `maxclust`, a warning will encourage you to increase your set `maxclust`.

The output of `clusso()` is assigned to the object `resreal`. `resreal` is a large list of lists. The recommended way to explore the results is to explore the `names(resreal)` and select each sub-list of interest.

```{r}
head(jbc)

system.time(resreal <- clusso(jbc, expected = expdeath, observed=death,timeperiod = period,
                              covars=FALSE, id = id, x= x,y = y, rMax =  rMax, 
                              utm=TRUE, analysis="both", model="poisson",maxclust=11))
```

We can use the function `clussopretty()` to create a nice table for the number of clusters detected:

```{r}
clussopretty(resreal, analysis="both", model="poisson",clusteridentify=FALSE)
```


With the argument `clusteridentify=TRUE`, centroid indices (by time period) are extracted. Since these lists are very long, output is left to the user to explore:

```{r, eval=FALSE}
#identify cluster IDs
clussopretty(resreal, analysis="both", model="poisson",clusteridentify=TRUE)
#identify cluster IDs where RR is greater or equal to 1.05
clussopretty(resreal, analysis="both", model="poisson", clusteridentify=TRUE, clusterRR = 1.05)
#identify cluster IDS that differ from the background rate
clussopretty(resreal, analysis="both", model="poisson", clusteridentify=TRUE, clusterRR = "background")
```


We can plot the output using `clussoplot()` to visualize the coefficient paths, where each line is the path for a potential cluster and the final selected clusters are identified with dashed vertical lines:

```{r}
clussoplot(resreal, analysis="both",model="poisson", Time=5, cv=FALSE)
```




## Comparison to Cross-Validation

We continue with the same data set as above, but now explore results by $k$-fold cross-validation. All initial inputs are assumed to be the same. We specify a total of 5 folds using the argument `cv=5` in the `clusso()` function.

```{r}
#perform clusso using cv (5 folds)
system.time(rescv <- clusso(jbc, expected = expdeath, observed=death,timeperiod = period, 
                            id=id, covars=FALSE, x= x,y = y, rMax =  rMax, utm=TRUE, 
                            analysis="both", model="poisson",maxclust=11, cv=5))

```

```{r}
clussopretty(rescv, analysis="both", model="poisson",clusteridentify=FALSE, cv=TRUE)
```




## Example with Covariates

We continue with the Japanese breast cancer data set imported above, but now consider covariates in the model that will be unpenalized. 

We explore the first 6 observations of the `jbc` dataframe using the `head()` function.



```{r}
#Set-up
head(jbc)
```


There are 5 additional covariates, `covar1` - `covar5` that we will want to adjust for in our Poisson or quasi-Poisson regression as unpenalized terms.


We now set the argument `covars=TRUE` to tell `clusso()` to identify the covariates and to include them as unpenalized terms:

```{r}
system.time(rescovars<- clusso(df=jbc, expected = expdeath, observed=death,
                               timeperiod = period, id=id, covars=TRUE,
                               x= x,y = y, rMax =  rMax, utm=TRUE, analysis="both", 
                               model="poisson",maxclust=11,overdispfloor = TRUE,
                               cv=NULL, collapsetime = FALSE))

```

Use `clussopretty()` to create a nice table of clusters detected:

```{r}
clussopretty(rescovars, analysis="both", model="poisson",clusteridentify = FALSE)
```

We can also extract the coefficients for the unpenalized terms and undo the log-link by exponentiating:

```{r}
out <- clussopretty(rescovars, analysis="both", model="poisson",covars=TRUE)
out$table.coefs
out$table.expcoefs
out$table.clusters
```




