---
title: "Using *clusso* with Maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to clusso I}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 5, fig.width = 8, message = FALSE, warning = FALSE)
```

#Introduction
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This package is still under development. Please [report any bugs](https://github.com/mkamenet3/clusso/issues).

Messages from `clusso` have been suppressed for this tutorial, but there are many messages generated that we hope the user will find helpful. Feedback is welcome [here](https://github.com/mkamenet3/clusso/issues).

This example uses simulated data which are meant for illustration of how to use `clusso` with maps in `R`.

##Preparing the Map

Load `clusso`, as well as the `tidyverse`, `tigris`, `sf`, `tmap`, and `sp` packages:

```{r, message=FALSE, warning=FALSE}
library("clusso")
library(tidyverse)
library(tigris)
library(sf)
library(tmap)
library(sp)

```

Using the `counties()` function from the  `tigris`, we load a county-level map of Wisconsin. We `fortify()` the map and create `wi_map` in order to use it with `ggplot2`:

```{r}
wi <- tigris::counties("Wisconsin", cb=TRUE)
wi_map <- fortify(wi)
```


To plot the counties of Wisconsin, we use `geom_map()` (part of `ggplot2`).

```{r}

p1 <- ggplot() + geom_map(data=wi_map, map=wi_map, 
                    aes(x=long, y=lat, map_id=id),
                    color="black", fill="white", size=0.25) +
    coord_map() +
    ylab("Latitude") + 
    xlab("Longitude")
p1
```

##Preparing the Data

Next, we simulate some data. For this example, we imagine that in each county there were a total number of people sampled (**ntrials**), some of which have outcome *X* (**ncases**). We assume the total number of people sampled come from a negative binomial distribution and use `rnbinom()` to simulate **ntrials**. We assume cases follow a binomial distribution and use `rbinom()` to draw cases from the population with probability $\pi$. For illustration of `clusso`, we put a cluster with elevated rates in Wood, Portage, Waushera, Adams, Juneau, and Marquette counties. We will only be looking at space for this example so there is only a single time period.


```{r}
#create ncases and ntrials for 72 counties in Wisconsin
set.seed(2)
ncounties <- length(unique(wi@data$NAME))
ntrials <- rnbinom(n=ncounties, mu=1000, size =10)
dat <- cbind.data.frame(id=unique(wi@data$NAME),ntrials)
dat <- dat %>%
    mutate(ncases = if_else(id %in% c("Wood", "Portage", "Waushara", "Adams", "Juneau","Marquette"), 
                            rbinom(ncounties,ntrials, 0.9), 
                            rbinom(ncounties,ntrials, 0.1)),
           time = "Period1")
str(dat)
```

The dataframe `dat`, that contains the county name (*id*), number of people sampled (*ntrials*), number of people with outcome *X* (*ncases*), and number of time periods (*time*). Next, we need to merge `dat` with the `SpatialPolygonsDataFrame`, `wi`.  We use `merge()` from the `sp` package to combine the two by county name (county name is the variable *NAME* in `wi` and is *id* in `dat`).

```{r}

#merge to map
widat <-  sp::merge(wi,dat, by.x="NAME", by.y="id", duplicateGeoms=TRUE)
class(widat)
head(widat)
```

In order to extract the centroid from each county, we first convert our new `SpatialPolygonsDataFrame` to a simple features (*sf*) object using `st_as_sf()` from the `sf` package. It is possible to extract centroids from the `SpatialPolygonsDataFrame`, but we do so here in order to also illustrate simple features.

```{r}
#convert to simple features sf
widatsf <- st_as_sf(widat)
class(widatsf)
```

We extract the centroids using `st_centroid()` and extract the coordinates using `st_coordinates()`; both functions are from the `sf` package.

```{r}
coords <- as.data.frame(st_coordinates(st_centroid(widatsf)))
head(coords)
```

We can plot the centroids on top of the base map we created above, *p1*:

```{r}
p1 + geom_point(data=coords,aes(x=X, y=Y), color="red", shape=3)

```

##Using `clusso`

We first set global parameters, namely the maximum radius size for each potential cluster (`rMax <- 100`), the number of time periods to 1 (`Time <- 1`). We also extract our *x* and *y* coordinates, which are in longitude and latitude, and set them as `xx` and `yy`.


To perform the LASSO-based cluster detection, we use the function `clusso()`. In the first argument, we set the name of the dataframe, `df=widatsf`. In the second argument we set `expected` counts equal to the total number of trials in the `widatsf` dataframe (`expected = ntrials`). In the third argument, we set the observed counts equal to the total number of cases in each county in the `widatsf` dataframe (`observed = ncases`). In the fourth argument, we set the timeperiod equal to the name of the time periods in the `widatsf` dataframe (`timeperiod = time`).  In this example, we do not want to include other covariates as un-penalized terms in the model so we set `covars = FALSE`. 

The next two arguments are `x` and `y`, which are the longitude and latitude for each centroids. Since these are longitude and latitude coordinates, we also set `utm=FALSE`. This will tell `clusso` to use the *great-circle-distance* or *Haversine* distance to calculate the distance between a centroid up to `rMax`.

The next argument is the maximum radius (in km) for the potential clusters. We set `rMax = rMax`, which we defined above as 100km. The argument `utm` specifies that our coordinates are *NOT* in UTM coordinates (`utm = FALSE`).  The is `analysis` argument defines the type of analysis to be performed. Options include `space`, `spacetime`, or `both`. Here, we set  `analysis` equal to `space`, which will perform only the spatial analysis. The argument `model` is the type of model to be used. In this example, we specify the *binomial* model. The argument `maxclust` is the maximum number of actual clusters expected in the given area. This is based on scientific knowledge of the area. For this dataset, we set `maxclust=15`. If the number of identified clusters exceeds `maxclust`, a warning will encourage you to increase your set `maxclust`.

The output of `clusso()` is assigned to the object `resclusso`. `resclusso` is a large list of lists. The recommended way to explore the results is to explore the `names(resreal)` and select each sub-list of interest.


```{r}
rMax <- 100
Time <- 1

xx <- coords$X
yy <- coords$Y


resclusso <- clusso(df=widatsf,
                    expected = ntrials,
                    observed = ncases,
                    timeperiod = time,
                    id = NAME,
                    covars = FALSE,
                    x= xx,
                    y= yy,
                    rMax = rMax,
                    utm=FALSE,
                    analysis = "space",
                    model="binomial",
                    maxclust = 15, collapsetime = TRUE)
clussopretty(resclusso, analysis="space", model="binomial",clusteridentify=FALSE)

```


##Plotting `clusso` Results on the Map

We set the max and min relative risk for the color scheme using `maxrr` and `minrr`. You  may want to explore these values depending on the presentation you are using (a shortened maxrr, minrr distance will bring colors in closer and maps on a projector will appear more vivid and darker).

We extract the relative risk estimates using BIC from `resclusso$lassoresult.p.s$E.qbic` and then use `redblue()` to map the estimated relative risks to colors.


```{r}
maxrr <- 2
minrr <- 0.5
estrr.bic <- resclusso$lassoresult.p.s$E.qbic
cluster_ix <- redblue(log(maxrr *  pmax(minrr, pmin(estrr.bic, maxrr)))/log(maxrr^2))
table(cluster_ix)
```

Next, we combine the estimated relative risks (`resclusso$lassoresult.p.s$E.qbic`), county names (widatsf$NAME), and mapped colors (`cluster_ix`) into the dataframe `rr.bic`. We map the county names onto the colors (`names(cluster_ix) <- rr.bic\$county`).

```{r}
rr.bic <- cbind.data.frame(rr=resclusso$lassoresult.p.s$E.qbic, 
                           county=widatsf$NAME, 
                           colors = cluster_ix)
names(cluster_ix) <- rr.bic$county
```

We merge the dataframe `rr.bic` to the `SpatialPolygonsDataFrame`, `wi`, by county name and convert it to a simple features dataframe for use with `ggplot2`.

```{r}
merged.bic<- sp::merge(wi,rr.bic, by.x="NAME", by.y="county", duplicateGeoms=TRUE)
datdf.bic <- st_as_sf(merged.bic)

```

Finally, we use `geom_sf()` to plot the estimated relative risks onto the map. We use the aesthetic (`aes(fill=NAME)`) and `scale_fill_manual(values=cluster_ix)` to plot our pre-defined colors onto the map.

```{r}

p1+ geom_sf(data=datdf.bic, aes(fill=NAME)) +
    scale_fill_manual(values=cluster_ix)+
    theme(legend.position = "none",
          plot.title = element_text(hjust=0.5)) +
    ggtitle("Binomial, Space: BIC") 

```

For a simpler map, we can also isolate the counties with relative risks greater than 1 and only outline county edges with red:

```{r}
ggplot()+ geom_sf(data=datdf.bic, lwd=0, fill='transparent') +
    geom_sf(data=subset(datdf.bic, rr>1), color="red") +
    theme_bw()+
    theme(legend.position = "none",
          plot.title = element_text(hjust=0.5)) +
    ggtitle("Binomial, Space: BIC") 
```

